---
title: "Hotels and Booking Cancellations"
author: "Ana Imam, Jihan Khandaker, Supreet Nirman, Al Shumyatcher, Mihir Zaveri"
date: "February 21st, 2021"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

## Introduction

In this project, we are utilizing the data set of hotel bookings to address the problem of cancellations. This is an important issue for hotels from a business standpoint, as cancellations result in a loss of potential revenue and limit the optimization of the hotel's service levels. Our analyses of this data set are aimed to help the management teams of both the City Hotel and Resort Hotel to better predict the likelihood of cancellation. With this, the hotels can then employ an overbooking method that can optimize the capacity-based revenue management. 

#### Approach

In approaching this data set, we will first separate the data for the City Hotel and the Resort Hotel. For both hotels, we have narrowed down a number of booking features that we believe have a significant correlation with the probability of cancellation. From this, we have developed a profile of probabilities that can be applied to any future bookings. Hotels can then use this profile to calculate the cancellation probability of future bookings and consider the option of overbooking. 

#### Data Set-up
```{r}
hotels <- read.csv("hotel_bookings.csv")
rhotels <- hotels[hotels$hotel == "Resort Hotel",]
chotels <- hotels[hotels$hotel == "City Hotel",]
library(ggplot2)
# install.packages(gridExtra)
library(gridExtra)
```

## Variable  Omission

  This data set provides information on many of the different features of hotel bookings. However, we have realized that not every variable has the potential of predicting cancellation probabilities, and thus, have been omitted:

**Arrival_date_year**: Since these are the years of the recorded bookings, they are not predictive of bookings in future years

**Arrival_date_week_number** & **Arrival_date_day_of_month**: These variables both have excessive variation and are also unnecessary given the analysis of the arrival months of the bookings

**adults**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$adults <- as.factor(rhotels$adults)
sort(summary(rhotels$adults), decreasing = TRUE) [1:5]
```
```{r}
chotels$adults <- as.factor(chotels$adults)
sort(summary(chotels$adults), decreasing = TRUE) [1:5]
```

**meal**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$meal <- as.factor(rhotels$meal)
sort(summary(rhotels$meal), decreasing = TRUE) [1:5]
```
```{r}
chotels$meal <- as.factor(chotels$meal)
sort(summary(chotels$meal), decreasing = TRUE) [1:5]
```

**distribution_channel**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$distribution_channel <- as.factor(rhotels$distribution_channel)
sort(summary(rhotels$distribution_channel), decreasing = TRUE) [1:5]
```
```{r}
chotels$distribution_channel <- as.factor(chotels$distribution_channel)
sort(summary(chotels$distribution_channel), decreasing = TRUE) [1:5]
```

**is_repeated_guest**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$is_repeated_guest <- as.factor(rhotels$is_repeated_guest)
sort(summary(rhotels$is_repeated_guest), decreasing = TRUE) [1:5]
```
```{r}
chotels$is_repeated_guest <- as.factor(chotels$is_repeated_guest)
sort(summary(chotels$is_repeated_guest), decreasing = TRUE) [1:5]
```

**previous_cancellations**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$previous_cancellations <- as.factor(rhotels$previous_cancellations)
sort(summary(rhotels$previous_cancellations), decreasing = TRUE) [1:5]
```
```{r}
chotels$previous_cancellations <- as.factor(chotels$previous_cancellations)
sort(summary(chotels$previous_cancellations), decreasing = TRUE) [1:5]
```

**previous_bookings_not_canceled**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$previous_bookings_not_canceled <- as.factor(rhotels$previous_bookings_not_canceled)
sort(summary(rhotels$previous_bookings_not_canceled), decreasing = TRUE) [1:5]
```
```{r}
chotels$previous_bookings_not_canceled <- as.factor(chotels$previous_bookings_not_canceled)
sort(summary(chotels$previous_bookings_not_canceled), decreasing = TRUE) [1:5]
```

**reserved_room_type**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$reserved_room_type <- as.factor(rhotels$reserved_room_type)
sort(summary(rhotels$reserved_room_type), decreasing = TRUE) [1:5]
```
```{r}
chotels$reserved_room_type <- as.factor(chotels$reserved_room_type)
sort(summary(chotels$reserved_room_type), decreasing = TRUE) [1:5]
```

**assigned_room_type**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$assigned_room_type <- as.factor(rhotels$assigned_room_type)
sort(summary(rhotels$assigned_room_type), decreasing = TRUE) [1:5]
```
```{r}
chotels$assigned_room_type <- as.factor(chotels$assigned_room_type)
sort(summary(chotels$assigned_room_type), decreasing = TRUE) [1:5]
```

**booking_changes**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$booking_changes <- as.factor(rhotels$booking_changes)
sort(summary(rhotels$booking_changes), decreasing = TRUE) [1:5]
```
```{r}
chotels$booking_changes <- as.factor(chotels$booking_changes)
sort(summary(chotels$booking_changes), decreasing = TRUE) [1:5]
```

**deposit_type**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$deposit_type <- as.factor(rhotels$deposit_type)
sort(summary(rhotels$deposit_type), decreasing = TRUE) [1:5]
```
```{r}
chotels$deposit_type <- as.factor(chotels$deposit_type)
sort(summary(chotels$deposit_type), decreasing = TRUE) [1:5]
```

**company**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$company <- as.factor(rhotels$company)
sort(summary(rhotels$company), decreasing = TRUE) [1:5]
```
```{r}
chotels$company <- as.factor(chotels$company)
sort(summary(chotels$company), decreasing = TRUE) [1:5]
```

**days_in_waiting_list**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$days_in_waiting_list <- as.factor(rhotels$days_in_waiting_list)
sort(summary(rhotels$days_in_waiting_list), decreasing = TRUE) [1:5]
```
```{r}
chotels$days_in_waiting_list <- as.factor(chotels$days_in_waiting_list)
sort(summary(chotels$days_in_waiting_list), decreasing = TRUE) [1:5]
```

**customer_type**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$customer_type <- as.factor(rhotels$customer_type)
sort(summary(rhotels$customer_type), decreasing = TRUE) [1:5]
```
```{r}
chotels$customer_type <- as.factor(chotels$customer_type)
sort(summary(chotels$customer_type), decreasing = TRUE) [1:5]
```

**adr**: since this feature records, lodging transactions rates, it is irrelevant to cancellations

**required_car_parking_spaces**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$required_car_parking_spaces <- as.factor(rhotels$required_car_parking_spaces)
sort(summary(rhotels$required_car_parking_spaces), decreasing = TRUE) [1:5]
```
```{r}
chotels$required_car_parking_spaces <- as.factor(chotels$required_car_parking_spaces)
sort(summary(chotels$required_car_parking_spaces), decreasing = TRUE) [1:5]
```




## Resort Hotel 

### Variable Analyses

**Lead_time**
```{r}
r_leadtime_cancel <- tapply(rhotels$is_canceled, rhotels$lead_time, mean, na.rm = TRUE)
plot(r_leadtime_cancel, xlab = "Lead Time in days")
```
As shown in this scatter plot, there is a slight increase in the cancellation probability as the lead time increases. However, this trend also comes along with increasing variation in the cancellation probabilities, as shown by the spread of points. The least amount of variation is under 100 days of lead time. So, we can assume that, in general, a greater lead time has a greater probability of cancellation, but especially if it is under 100 days.

**arrival_date_month**
```{r}
rhotels$arrival_date_month <- as.factor(rhotels$arrival_date_month)
sort(summary(rhotels$arrival_date_month), decreasing = TRUE)
```
*Variable Validation: since this variable has enough variation and each month has a significant sample size, we can proceed to analyzing the cancellation rates without omitting any observations
```{r}
rcancelledbymonth <- tapply(rhotels$is_canceled, rhotels$arrival_date_month, mean, na.rm = TRUE)
sort(rcancelledbymonth, decreasing = TRUE)
```
As shown in the table, August and June have the highest cancellation rates of about 0.33, which is about double the cancellation rates of November and January. Given this variation in cancellation rates, the hotel should definitely consider the month when calculating the probability of cancellation for each booking.

**Stays_in_weekend_nights**
```{r}
r_weekend_cancel <- tapply(rhotels$is_canceled, rhotels$stays_in_weekend_nights, mean, na.rm = TRUE)
plot(r_weekend_cancel, xlab = "Weekend Days Booked")
```
As shown in the scatter plot, the cancellation rates gradually increases to about 0.6 until about 10 weekend days booked. After that, the cancellation rate drops to nearly 0. Given this, the hotel can assume that if there are under 10 weekend days booked, the probability of cancellation will gradually increase up to 0.6 at most. 

**Stays_in_week_nights**
```{r}
r_weekday_cancel <- tapply(rhotels$is_canceled, rhotels$stays_in_week_nights, mean, na.rm = TRUE)
plot(r_weekday_cancel, xlab = "Weekday Nights Booked")
```
As shown in the scatter plot, the cancellation rates are generally consistent around 0.3 until about 10 weekday nights booked. After that, the cancellation rate varies greatly. Given this inconsistent variation in cancellation rates, the hotel does not need to consider the number of weekday nights booked as much when calculating the overall probability of cancellation for a booking.

**children**: children=1 or not=0
```{r}
rhotels$children2 <- ifelse(rhotels$children > 0, 1, 0)
rhotels$children2 <- as.factor(rhotels$children2)
r_cancel_children <- tapply(rhotels$is_canceled, rhotels$children2, mean, na.rm = TRUE)
r_cancel_children
```

**babies**: babies=1 or not=0
```{r}
rhotels$babies2 <- ifelse(rhotels$babies > 0, 1, 0)
rhotels$babies2 <- as.factor(rhotels$babies2)
r_cancel_babies <- tapply(rhotels$is_canceled, rhotels$babies2, mean, na.rm = TRUE)
r_cancel_babies
```

**country** <-- how to filter top 10 or with > 100 bookings? tapply sum or mean?
```{r}
rhotels$country <- as.factor(rhotels$country)
nlevels(rhotels$country)
sort(summary(rhotels$country), decreasing = TRUE)
```

```{r}
rcancelledbycountry <- tapply(rhotels$is_canceled, rhotels$country, mean, na.rm = TRUE)
sort(rcancelledbycountry, decreasing = TRUE)
```

**market_segment**
```{r}
rhotels$market_segment <- as.factor(rhotels$market_segment)
nlevels(rhotels$market_segment)
sort(summary(rhotels$market_segment), decreasing = TRUE)
rcancelledbymarket <- tapply(rhotels$is_canceled, rhotels$market_segment, mean, na.rm = TRUE)
sort(rcancelledbymarket, decreasing = TRUE)
```

**agent** <-- how to filter top 10 or with > 100 bookings?
```{r}
rhotels$agent <- as.factor(rhotels$agent)
nlevels(rhotels$agent)
sort(summary(rhotels$agent), decreasing = TRUE) [1:10]
rcancelledbyagent <- tapply(rhotels$is_canceled, rhotels$agent, mean, na.rm = TRUE)
sort(rcancelledbyagent, decreasing = TRUE) [1:10]
```

**total_of_special_requests** <-- how to filter with > 100 bookings?
```{r}
rhotels$total_of_special_requests <- as.factor(rhotels$total_of_special_requests)
nlevels(rhotels$total_of_special_requests)
sort(summary(rhotels$total_of_special_requests), decreasing = TRUE)
rcancelledbyrequest <- tapply(rhotels$is_canceled, rhotels$total_of_special_requests, mean, na.rm = TRUE)
sort(rcancelledbyrequest, decreasing = TRUE)
```

### Booking Profile

## City Hotel 

### Variable Analyses

**Lead_time**
```{r}
c_leadtime_cancel <- tapply(chotels$is_canceled, chotels$lead_time, mean, na.rm = TRUE)
plot(c_leadtime_cancel, xlab = "Lead Time in days")
```
As shown in this scatter plot, there is a significant increase in the cancellation probability as the lead time increases. Though there is some variation, the overall trend is quite strong. Overall, we can assume that a greater lead time has a greater probability of cancellation. 

**arrival_date_month**
```{r}
chotels$arrival_date_month <- as.factor(chotels$arrival_date_month)
sort(summary(chotels$arrival_date_month), decreasing = TRUE)
```
*Variable Validation: since this variable has enough variation and each month has a significant sample size, we can proceed to analyzing the cancellation rates without omitting any observations
```{r}
ccancelledbymonth <- tapply(chotels$is_canceled, chotels$arrival_date_month, mean, na.rm = TRUE)
sort(ccancelledbymonth, decreasing = TRUE)
```
As shown in the table, April has the highest cancellation rate of about 0.46 while March has the lowest cancellation rate of about 0.37, which is only a 0.10 difference. Given that the cancellation rates do not vary greatly between any of the months, the hotel does not need to consider the month of a booking as much when calculating its overall probability of cancellation.

**Stays_in_weekend_nights**
```{r}
c_weekend_cancel <- tapply(chotels$is_canceled, chotels$stays_in_weekend_nights, mean, na.rm = TRUE)
plot(c_weekend_cancel, xlab = "Weekend Days Booked")
```
As shown in the scatter plot, the cancellation rates gradually increases to about 0.8 until about 11 weekend days booked. After that, the cancellation rate drops to nearly 0. Given this, the hotel can assume that if there are under 11 weekend days booked, the probability of cancellation will gradually increase up to 0.8 at most. 

**Stays_in_week_nights**
```{r}
c_weekday_cancel <- tapply(chotels$is_canceled, chotels$stays_in_week_nights, mean, na.rm = TRUE)
plot(c_weekday_cancel, xlab = "Weekday Nights Booked")
```
As shown in the scatter plot, the cancellation rates increases until about 25 weekend nights booked. After that, the cancellation rate drops to nearly 0. Given this, the hotel can assume that if there are under 25 weekday nights booked, the probability of cancellation will gradually increase. 

**children**: children=1 or not=0
```{r}
chotels$children2 <- ifelse(chotels$children > 0, 1, 0)
chotels$children2 <- as.factor(chotels$children2)
c_cancel_children <- tapply(chotels$is_canceled, chotels$children2, mean, na.rm = TRUE)
c_cancel_children
```

**babies**: babies=1 or not=0
```{r}
chotels$babies2 <- ifelse(chotels$babies > 0, 1, 0)
chotels$babies2 <- as.factor(chotels$babies2)
c_cancel_babies <- tapply(chotels$is_canceled, chotels$babies2, mean, na.rm = TRUE)
c_cancel_babies
```

**country** <-- how to filter top 10 or with > 100 bookings? tapply sum or mean?
```{r}
chotels$country <- as.factor(chotels$country)
nlevels(chotels$country)
sort(summary(chotels$country), decreasing = TRUE)
ccancelledbycountry <- tapply(chotels$is_canceled, chotels$country, mean, na.rm = TRUE)
sort(ccancelledbycountry, decreasing = TRUE)
```

**market_segment** <-- how to filter with > 100 bookings?
```{r}
chotels$market_segment <- as.factor(chotels$market_segment)
nlevels(chotels$market_segment)
sort(summary(chotels$market_segment), decreasing = TRUE) 
ccancelledbymarket <- tapply(chotels$is_canceled, chotels$market_segment, mean, na.rm = TRUE)
sort(ccancelledbymarket, decreasing = TRUE) 
```

**agent** <-- how to filter top 10 or with > 100 bookings?
```{r}
chotels$agent <- as.factor(chotels$agent)
nlevels(chotels$agent)
sort(summary(chotels$agent), decreasing = TRUE) [1:10]
ccancelledbyagent <- tapply(chotels$is_canceled, chotels$agent, mean, na.rm = TRUE)
sort(ccancelledbyagent, decreasing = TRUE) [1:10]
```

**total_of_special_requests** <-- how to filter with > 100 bookings?
```{r}
chotels$total_of_special_requests <- as.factor(chotels$total_of_special_requests)
nlevels(chotels$total_of_special_requests)
sort(summary(chotels$total_of_special_requests), decreasing = TRUE)
ccancelledbyrequest <- tapply(chotels$is_canceled, chotels$total_of_special_requests, mean, na.rm = TRUE)
sort(ccancelledbyrequest, decreasing = TRUE)
```

**cancellations by number of adults and children**
```{r, fig.width = 15}
#, fig.height = 5}
colorPlot <- function(name, h = hotels){

  ifelse(name == "City Hotel", x <- chotels, x<- rhotels)
   #      ifelse(thing == "Airline", x <- n$carrier, x <- n$timeofday))
  
  a <- as.data.frame( aggregate(is_canceled ~ adults + children, data = x, mean, na.rm = TRUE) )
  
  SEQ <- function(x){
    return (seq(0, x, by = 1))
  }
  
  qplot(a$adults,
        a$children,
        color = a$is_canceled,
        size = I(10),
        geom = "point",
        alpha = I(1),
        xlab = "Adults",
        ylab = "Children") +
    
    scale_color_gradient(
      limits = c(0, max(a$is_canceled)),
      low = "green", high = "red",
      name = "Cancellation Probability") +
    
    ggtitle(paste("Cancellation Probability as a Function of Number of\n Adults and Children for", name)) +
    
    theme( plot.title = element_text( hjust = 0.5 ) ) +
    theme( panel.grid.minor = element_blank() ) +
    theme_bw()
    
    #scale_x_continuous(labels = SEQ(6), breaks = SEQ(6)) +
    #scale_y_continuous(labels = SEQ(8), breaks = SEQ(8))
    
}

plot1 <- colorPlot(name = "City Hotel")
plot2 <- colorPlot(name = "Resort Hotel")
grid.arrange(plot1, plot2, ncol=2)

#par(mfrow = c(1,2))



```

### Booking Profile

## Conclusion

