---
title: "Hotels and Booking Cancellations"
author: "Ana Imam, Jihan Khandaker, Supreet Nirman, Al Shumyatcher, Mihir Zaveri"
date: "February 21st, 2021"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

## Introduction

In this project, we are utilizing the data set of hotel bookings to address the problem of cancellations. This is an important issue for hotels from a business standpoint, as cancellations result in a loss of potential revenue and limit the optimization of the hotel's service levels. Our analyses of this data set are aimed to help the management teams of both the City Hotel and Resort Hotel to better predict the likelihood of cancellation. With this, the hotels can then employ an overbooking method that can optimize the capacity-based revenue management. 

#### Approach

In approaching this data set, we will first separate the data for the City Hotel and the Resort Hotel. For both hotels, we have narrowed down a number of booking features that we believe have a significant correlation with the probability of cancellation. From this, we have developed a profile of probabilities that can be applied to any future bookings. Hotels can then use this profile to calculate the cancellation probability of future bookings and consider the option of overbooking. 

#### Data Set-up
```{r}
hotels <- read.csv("hotel_bookings.csv")
rhotels <- hotels[hotels$hotel == "Resort Hotel",]
chotels <- hotels[hotels$hotel == "City Hotel",]
library(ggplot2)
# install.packages("gridExtra")
library(gridExtra)
```
#### Graph Set-up and Design
```{r}
weekend_cancellations_by_hotel <- function(name = "HOTEL"){
  ifelse(name == "City", hot <- chotels, hot <- rhotels)
  a <- as.data.frame( table( hot$stays_in_weekend_nights, hot$is_canceled ) )
  a <- as.data.frame( cbind.data.frame( a[1:(nrow(a)/2),1],
                                        a[1:(nrow(a)/2),3],
                                        a[(nrow(a)/2+1):nrow(a),3] )
                      )
  colnames(a) <- c("Weekends", "NotCanceled", "Canceled")
  a$percentCanceled <- a$Canceled / (a$Canceled + a$NotCanceled)
  a <- a[a$Canceled + a$NotCanceled >= 10, ]
  a <- tapply( a$percentCanceled, a$Weekends, mean, na.rm = TRUE )
  #the pretty colors serve as but a temporary repreive from the torture that is trying to write code in R
  barplot( a[!is.na(a)] , xlab = "Weekend Nights", ylab = "Cancellation Probability", main = paste(name, "Hotel"), axis.lty=1, ylim = c(0,0.8), col = rgb(255 * a[!is.na(a)], 225 * (1-a[!is.na(a)]), 0, maxColorValue = 255))
  
  
}
```
```{r}
weeknight_cancellations_by_hotel <- function(name = "HOTEL"){
  ifelse(name == "City", hot <- chotels, hot <- rhotels)
  a <- as.data.frame( table( hot$stays_in_week_nights, hot$is_canceled ) )
  a <- as.data.frame( cbind.data.frame( a[1:(nrow(a)/2),1],
                                        a[1:(nrow(a)/2),3],
                                        a[(nrow(a)/2+1):nrow(a),3] )
                      )
  colnames(a) <- c("Weeknights", "NotCanceled", "Canceled")
  a$percentCanceled <- a$Canceled / (a$Canceled + a$NotCanceled)
  a <- a[a$Canceled + a$NotCanceled >= 10, ]
  a <- tapply( a$percentCanceled, a$Weeknights, mean, na.rm = TRUE )
  barplot( a[!is.na(a)] , xlab = "Week Nights", ylab = "Cancellation Probability", main = paste(name, "Hotel"), axis.lty=1, ylim = c(0,0.8), col = rgb(255 * a[!is.na(a)], 225 * (1-a[!is.na(a)]), 0, maxColorValue = 255))
  
  
}
```
```{r}
week_end_colorPlot <- function(name, x = hotels, probabilityLimit = 9001, minNumber = 30){
  
  ifelse(name == "City Hotel", x <- chotels, x<- rhotels)
  x$ones <- 1
  
  a <- aggregate(is_canceled ~ stays_in_week_nights + stays_in_weekend_nights, data = x, mean, na.rm = TRUE)
  onesFrame <- aggregate(ones ~ stays_in_week_nights + stays_in_weekend_nights, data = x, sum, na.rm = TRUE)
  
  filteredA <- cbind(a[,1][onesFrame$ones > minNumber][a$is_canceled < probabilityLimit],
                     a[,2][onesFrame$ones > minNumber][a$is_canceled < probabilityLimit],
                     a[,3][onesFrame$ones > minNumber][a$is_canceled < probabilityLimit]
                     )
  
  colnames(filteredA) <- c("stays_in_week_nights", "stays_in_weekend_nights", "Canceled")
  filteredA <- as.data.frame( filteredA[complete.cases(filteredA), ] )
  
  qplot(filteredA$stays_in_week_nights,
        filteredA$stays_in_weekend_nights,
        color = filteredA$Canceled,
        size = I(12),
        geom = "point",
        alpha = I(0.85),
        xlab = "Stays in Week Nights",
        ylab = "Stays in Weekend Nights") +
    
    scale_color_gradient(
      limits = c(0, ceiling( max(filteredA$Canceled) / 0.2) * 0.2),
      low = "green", high = "red",
      name = "Cancellation Probability") +
    
    ggtitle(paste("Cancellation Probability as a Function of Number\n of Week and Weekend Nights for", name)) +
    
    theme( plot.title = element_text( hjust = 0.5 ) ) +
    theme( panel.grid.minor = element_blank() ) +
    theme_bw()
    
}
```


## Variable  Omission

  This data set provides information on many of the different features of hotel bookings. However, we have realized that not every variable has the potential of predicting cancellation probabilities, and thus, have been omitted:

**Arrival_date_year**: Since these are the years of the recorded bookings, they are not predictive of bookings in future years

**Arrival_date_week_number** & **Arrival_date_day_of_month**: These variables both have excessive variation and are also unnecessary given the analysis of the arrival months of the bookings

**adults**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$adults <- as.factor(rhotels$adults)
sort(summary(rhotels$adults), decreasing = TRUE) [1:5]
```
```{r}
chotels$adults <- as.factor(chotels$adults)
sort(summary(chotels$adults), decreasing = TRUE) [1:5]
```

**meal**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$meal <- as.factor(rhotels$meal)
sort(summary(rhotels$meal), decreasing = TRUE) [1:5]
```
```{r}
chotels$meal <- as.factor(chotels$meal)
sort(summary(chotels$meal), decreasing = TRUE) [1:5]
```

**distribution_channel**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$distribution_channel <- as.factor(rhotels$distribution_channel)
sort(summary(rhotels$distribution_channel), decreasing = TRUE) [1:5]
```
```{r}
chotels$distribution_channel <- as.factor(chotels$distribution_channel)
sort(summary(chotels$distribution_channel), decreasing = TRUE) [1:5]
```

**is_repeated_guest**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$is_repeated_guest <- as.factor(rhotels$is_repeated_guest)
sort(summary(rhotels$is_repeated_guest), decreasing = TRUE) [1:5]
```
```{r}
chotels$is_repeated_guest <- as.factor(chotels$is_repeated_guest)
sort(summary(chotels$is_repeated_guest), decreasing = TRUE) [1:5]
```

**previous_cancellations**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$previous_cancellations <- as.factor(rhotels$previous_cancellations)
sort(summary(rhotels$previous_cancellations), decreasing = TRUE) [1:5]
```
```{r}
chotels$previous_cancellations <- as.factor(chotels$previous_cancellations)
sort(summary(chotels$previous_cancellations), decreasing = TRUE) [1:5]
```

**previous_bookings_not_canceled**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$previous_bookings_not_canceled <- as.factor(rhotels$previous_bookings_not_canceled)
sort(summary(rhotels$previous_bookings_not_canceled), decreasing = TRUE) [1:5]
```
```{r}
chotels$previous_bookings_not_canceled <- as.factor(chotels$previous_bookings_not_canceled)
sort(summary(chotels$previous_bookings_not_canceled), decreasing = TRUE) [1:5]
```

**reserved_room_type**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$reserved_room_type <- as.factor(rhotels$reserved_room_type)
sort(summary(rhotels$reserved_room_type), decreasing = TRUE) [1:5]
```
```{r}
chotels$reserved_room_type <- as.factor(chotels$reserved_room_type)
sort(summary(chotels$reserved_room_type), decreasing = TRUE) [1:5]
```

**assigned_room_type**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$assigned_room_type <- as.factor(rhotels$assigned_room_type)
sort(summary(rhotels$assigned_room_type), decreasing = TRUE) [1:5]
```
```{r}
chotels$assigned_room_type <- as.factor(chotels$assigned_room_type)
sort(summary(chotels$assigned_room_type), decreasing = TRUE) [1:5]
```

**booking_changes**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$booking_changes <- as.factor(rhotels$booking_changes)
sort(summary(rhotels$booking_changes), decreasing = TRUE) [1:5]
```
```{r}
chotels$booking_changes <- as.factor(chotels$booking_changes)
sort(summary(chotels$booking_changes), decreasing = TRUE) [1:5]
```

**deposit_type**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$deposit_type <- as.factor(rhotels$deposit_type)
sort(summary(rhotels$deposit_type), decreasing = TRUE) [1:5]
```
```{r}
chotels$deposit_type <- as.factor(chotels$deposit_type)
sort(summary(chotels$deposit_type), decreasing = TRUE) [1:5]
```

**company**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$company <- as.factor(rhotels$company)
sort(summary(rhotels$company), decreasing = TRUE) [1:5]
```
```{r}
chotels$company <- as.factor(chotels$company)
sort(summary(chotels$company), decreasing = TRUE) [1:5]
```

**days_in_waiting_list**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$days_in_waiting_list <- as.factor(rhotels$days_in_waiting_list)
sort(summary(rhotels$days_in_waiting_list), decreasing = TRUE) [1:5]
```
```{r}
chotels$days_in_waiting_list <- as.factor(chotels$days_in_waiting_list)
sort(summary(chotels$days_in_waiting_list), decreasing = TRUE) [1:5]
```

**customer_type**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$customer_type <- as.factor(rhotels$customer_type)
sort(summary(rhotels$customer_type), decreasing = TRUE) [1:5]
```
```{r}
chotels$customer_type <- as.factor(chotels$customer_type)
sort(summary(chotels$customer_type), decreasing = TRUE) [1:5]
```

**adr**: since this feature records, lodging transactions rates, it is irrelevant to cancellations

**required_car_parking_spaces**: as shown below, this feature lacks variability in both hotel populations, so it cannot be predictive of cancellations
```{r}
rhotels$required_car_parking_spaces <- as.factor(rhotels$required_car_parking_spaces)
sort(summary(rhotels$required_car_parking_spaces), decreasing = TRUE) [1:5]
```
```{r}
chotels$required_car_parking_spaces <- as.factor(chotels$required_car_parking_spaces)
sort(summary(chotels$required_car_parking_spaces), decreasing = TRUE) [1:5]
```

## Resort Hotel 

### **Lead_time**
```{r}
r_leadtime_cancel <- tapply(rhotels$is_canceled, rhotels$lead_time, mean, na.rm = TRUE)
plot(r_leadtime_cancel, xlab = "Lead Time in days")
```

As shown in this scatter plot, there is a slight increase in the cancellation probability as the lead time increases. However, this trend also comes along with increasing variation in the cancellation probabilities, as shown by the spread of points. The least amount of variation is under 100 days of lead time. So, we can assume that, in general, a greater lead time has a greater probability of cancellation, but especially if it is under 100 days.

### **arrival_date_month**
```{r}
rhotels$arrival_date_month <- as.factor(rhotels$arrival_date_month)
sort(summary(rhotels$arrival_date_month), decreasing = TRUE)
```

*Variable Validation: since this variable has enough variation and each month has a significant sample size, we can proceed to analyzing the cancellation rates without omitting any observations
```{r}
rcancelledbymonth <- tapply(rhotels$is_canceled, rhotels$arrival_date_month, mean, na.rm = TRUE)
sort(rcancelledbymonth, decreasing = TRUE)
```

As shown in the table, August and June have the highest cancellation rates of about 0.33, which is about double the cancellation rates of November and January. Given this variation in cancellation rates, the hotel should definitely consider the month when calculating the probability of cancellation for each booking.
```{r}
data <- data.frame(
  name=c("August (Highest Cancellation Month)","January (Lowest Cancellation Month") ,  
  value=c(0.3344912,0.1481988)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```

### **Stays_in_weekend_nights**
```{r}
r_weekend_cancel <- tapply(rhotels$is_canceled, rhotels$stays_in_weekend_nights, mean, na.rm = TRUE)
plot(r_weekend_cancel, xlab = "Weekend Nights Booked")
```

As shown in the scatter plot, the cancellation rates gradually increases to about 0.6 until about 10 weekend nights booked. After that, the cancellation rate drops to nearly 0. Given this, the hotel can assume that if there are under 10 weekend nights booked, the probability of cancellation will gradually increase up to 0.6 at most. 

The average cancel rates for each number of weekend nights booked is shown in the bar graph below:
```{r}
weekend_cancellations_by_hotel("Resort")
```

```{r}
data <- data.frame(
  name=c("7 Weekends (Highest Cancellation)","12 Weekends (Lowest Cancellation Month)") ,  
  value=c(0.6153846,0.2)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **Stays_in_week_nights**
```{r}
r_weekday_cancel <- tapply(rhotels$is_canceled, rhotels$stays_in_week_nights, mean, na.rm = TRUE)
plot(r_weekday_cancel, xlab = "Week Nights Booked")
```

As shown in the scatter plot, the cancellation rates are generally consistent around 0.3 until about 10 week nights booked. After that, the cancellation rate varies greatly. Given this inconsistent variation in cancellation rates, the hotel does not need to consider the number of week nights booked as much when calculating the overall probability of cancellation for a booking.

The average cancel rates for each number of week nights booked is shown in the bar graph below:
```{r}
weeknight_cancellations_by_hotel("Resort")
```

```{r}
data <- data.frame(
  name=c("22 Weekdays (Highest Cancellation)","0 Weekdays (Lowest Cancellation)") ,  
  value=c(1.0, 0.1614467)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **Stays in weekend and week nights**
Since many bookings include a combination of weekend and week nights, we have graphed the probability of cancellation by combining both variables:
```{r}
week_end_colorPlot(name = "Resort Hotel")#, minNumber = 5)
```

As shown here, longer stays have higher probabilities of cancellation.

### **children**
```{r}
rhotels$children2 <- ifelse(rhotels$children > 0, 1, 0)
rhotels$children2 <- as.factor(rhotels$children2)
r_cancel_children <- tapply(rhotels$is_canceled, rhotels$children2, mean, na.rm = TRUE)
r_cancel_children
```

The cancellation rate for bookings without children is about 0.27 while the cancellation rate for bookings with children is about 0.38. Given this fairly large difference, the hotel should consider if there are children when calculating the probability of cancellation for a booking.
```{r}
data <- data.frame(
  name=c("Have Children (Highest Cancellation)","No Children (Lowest Cancellation)") ,  
  value=c(0.3771527, 0.2681540)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **babies**
```{r}
rhotels$babies2 <- ifelse(rhotels$babies > 0, 1, 0)
rhotels$babies2 <- as.factor(rhotels$babies2)
r_cancel_babies <- tapply(rhotels$is_canceled, rhotels$babies2, mean, na.rm = TRUE)
r_cancel_babies
```

The cancellation rate for bookings without babies is about 0.28 while the cancellation rate for bookings with babies is about 0.19. Given this fairly large difference, the hotel should consider if there are babies when calculating the probability of cancellation for a booking.
```{r}
data <- data.frame(
  name=c("No Babies (Highest Cancellation)","Have Babies (Lowest Cancellation)") ,  
  value=c(0.278873, 0.1879562)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```

### **country** 
```{r}
rhotels$country <- as.factor(rhotels$country)
sort(summary(rhotels$country), decreasing = TRUE)
```

*Variable Validation: though this variable has enough variation, not every country has a significant sample size. Therefore, when analyzing the cancellation rates, we will omit countries with less than 1000 bookings.
```{r}
sort(summary(rhotels$country), decreasing = TRUE) [1:6]
rcancelledbycountry <- tapply(rhotels$is_canceled, rhotels$country, mean, na.rm = TRUE)
sort(rcancelledbycountry, decreasing = TRUE)
```

PRT, GBR, ESP, IRL, FRA, and DEU have cancellation rates of about 0.42, 0.13, 0.22, 0.20, 0.13, and 0.12, respectively. 

This variation shows that the cancellation rates are independent from the total number of bookings, so the country of the booking should be considered when calculating the probability of cancellation (if there has been at least 1000 bookings before).
```{r}
data <- data.frame(
  name=c("Portugal (Highest Cancellation)","LTU (Lowest Cancellation)") ,  
  value=c(0.42189450, 0.12136326)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```

### **market_segment**
```{r}
rhotels$market_segment <- as.factor(rhotels$market_segment)
sort(summary(rhotels$market_segment), decreasing = TRUE)
```

*Variable Validation: though this variable has enough variation, not every market segment has a significant sample size. Therefore, when analyzing the cancellation rates, we will omit market segments with less than 1000 bookings.
```{r}
sort(summary(rhotels$market_segment), decreasing = TRUE) [1:5]
rcancelledbymarket <- tapply(rhotels$is_canceled, rhotels$market_segment, mean, na.rm = TRUE)
sort(rcancelledbymarket, decreasing = TRUE)
```

As shown in the tables above, the Groups and Online TA market segments have significantly higher cancellation rates in comparison to other market segments. Given this variation, the hotel should consider the market segment when calculating the probability of cancellation for a booking (if there has been at least 1000 bookings before).
```{r}
data <- data.frame(
  name=c("Groups (Highest Cancellation)","Direct (Lowest Cancellation)") ,  
  value=c(0.4239205, 0.1348073)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **agent** 
```{r}
rhotels$agent <- as.factor(rhotels$agent)
sort(summary(rhotels$agent), decreasing = TRUE)
```

*Variable Validation: though this variable has enough variation, not every agent has a significant sample size. Therefore, when analyzing the cancellation rates, we will omit agents with less than 1000 bookings.
```{r}
sort(summary(rhotels$agent), decreasing = TRUE) [1:5]
rcancelledbyagent <- tapply(rhotels$is_canceled, rhotels$agent, mean, na.rm = TRUE)
sort(rcancelledbyagent, decreasing = TRUE)
```

Agents 240, 250, 241, and 40 have cancellation rates of about 0.39, 0.18, 0.14, and 0.08, respectively. 

This variation shows that the cancellation rates are independent from the total number of bookings, so the agent of the booking should be considered when calculating the probability of cancellation (if there has been at least 1000 bookings before).
```{r}
data <- data.frame(
  name=c("240 (Highest Cancellation)","28 (Lowest Cancellation)") ,  
  value=c(0.39431859, 0.08183633)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **total_of_special_requests**
```{r}
rhotels$total_of_special_requests <- as.factor(rhotels$total_of_special_requests)
sort(summary(rhotels$total_of_special_requests), decreasing = TRUE)
```

*Variable Validation: though this variable has enough variation, not every number has a significant sample size. Therefore, when analyzing the cancellation rates, we will omit numbers with less than 1000 bookings.
```{r}
sort(summary(rhotels$total_of_special_requests), decreasing = TRUE) [1:3]
rcancelledbyrequest <- tapply(rhotels$is_canceled, rhotels$total_of_special_requests, mean, na.rm = TRUE)
sort(rcancelledbyrequest, decreasing = TRUE)
```

As shown in the table above, as the number of special requests increase, the cancellation rate slightly decreases. However, since there is not much variation between the cancellation rates overall, this does not need to be considered when calculating the probability of cancellation for a booking.
```{r}
data <- data.frame(
  name=c("0 (Highest Cancellation)","5 (Lowest Cancellation)") ,  
  value=c(0.32270471, 0.07142857)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


## City Hotel 

### **Lead_time**
```{r}
c_leadtime_cancel <- tapply(chotels$is_canceled, chotels$lead_time, mean, na.rm = TRUE)
plot(c_leadtime_cancel, xlab = "Lead Time in days")
```

As shown in this scatter plot, there is a significant increase in the cancellation probability as the lead time increases. Though there is some variation, the overall trend is quite strong. Overall, we can assume that a greater lead time has a greater probability of cancellation. 

### **arrival_date_month**
```{r}
chotels$arrival_date_month <- as.factor(chotels$arrival_date_month)
sort(summary(chotels$arrival_date_month), decreasing = TRUE)
```

*Variable Validation: since this variable has enough variation and each month has a significant sample size, we can proceed to analyzing the cancellation rates without omitting any observations
```{r}
ccancelledbymonth <- tapply(chotels$is_canceled, chotels$arrival_date_month, mean, na.rm = TRUE)
sort(ccancelledbymonth, decreasing = TRUE)
```

As shown in the table, April has the highest cancellation rate of about 0.46 while March has the lowest cancellation rate of about 0.37, which is only a 0.10 difference. Given that the cancellation rates do not vary greatly between any of the months, the hotel does not need to consider the month of a booking as much when calculating its overall probability of cancellation.
```{r}
data <- data.frame(
  name=c("April (Highest Cancellation)","March (Lowest Cancellation)") ,  
  value=c(0.4632353, 0.3694642)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **Stays_in_weekend_nights**
```{r}
c_weekend_cancel <- tapply(chotels$is_canceled, chotels$stays_in_weekend_nights, mean, na.rm = TRUE)
plot(c_weekend_cancel, xlab = "Weekend Nights Booked")
```

As shown in the scatter plot, the cancellation rates gradually increases to about 0.8 until about 11 weekend nights booked. After that, the cancellation rate drops to nearly 0. Given this, the hotel can assume that if there are under 11 weekend nights booked, the probability of cancellation will gradually increase up to 0.8 at most. 

The average cancel rates for each number of weekend nights booked is shown in the bar graph below:
```{r}
weekend_cancellations_by_hotel("City")
```
```{r}
data <- data.frame(
  name=c("7 Stays (Highest Cancellation)","1 Stay (Lowest Cancellation)") ,  
  value=c(1, 0.3992255)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **Stays_in_week_nights**
```{r}
c_weekday_cancel <- tapply(chotels$is_canceled, chotels$stays_in_week_nights, mean, na.rm = TRUE)
plot(c_weekday_cancel, xlab = "Week Nights Booked")
```

As shown in the scatter plot, the cancellation rates increases until about 25 week nights booked. After that, the cancellation rate drops to nearly 0. Given this, the hotel can assume that if there are under 25 week nights booked, the probability of cancellation will gradually increase. 

The average cancel rates for each number of week nights booked is shown in the bar graph below:
```{r}
weeknight_cancellations_by_hotel("City")
```
```{r}
data <- data.frame(
  name=c("21 Stays (Highest Cancellation)","0 Stays (Lowest Cancellation)") ,  
  value=c(1, 0.2982067)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **Stays in weekend and week nights**
Since many bookings include a combination of weekend and week nights, we have graphed the probability of cancellation by combining both variables:
```{r}
week_end_colorPlot(name = "City Hotel")#, minNumber = 5)
```

As shown here, longer stays have higher probabilities of cancellation.

### **children**
```{r}
chotels$children2 <- ifelse(chotels$children > 0, 1, 0)
chotels$children2 <- as.factor(chotels$children2)
c_cancel_children <- tapply(chotels$is_canceled, chotels$children2, mean, na.rm = TRUE)
c_cancel_children
```

The cancellation rate for bookings without children is about 0.42 while the cancellation rate for bookings with children is about 0.35. Given this fairly large difference, the hotel should consider if there are children when calculating the probability of cancellation for a booking.
```{r}
data <- data.frame(
  name=c("No Children (Highest Cancellation)","Have Children (Lowest Cancellation)") ,  
  value=c(0.4215306, 0.3548766)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")
```

### **babies**
```{r}
chotels$babies2 <- ifelse(chotels$babies > 0, 1, 0)
chotels$babies2 <- as.factor(chotels$babies2)
c_cancel_babies <- tapply(chotels$is_canceled, chotels$babies2, mean, na.rm = TRUE)
c_cancel_babies
```

The cancellation rate for bookings without babies is about 0.42 while the cancellation rate for bookings with babies is about 0.17. Given this significantly large difference, the hotel should consider if there are babies when calculating the probability of cancellation for a booking.
```{r}
data <- data.frame(
  name=c("No Babies (Highest Cancellation)","Have Babies (Lowest Cancellation)") ,  
  value=c(0.4184091, 0.1734417)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **country**
```{r}
chotels$country <- as.factor(chotels$country)
sort(summary(chotels$country), decreasing = TRUE)
```

*Variable Validation: though this variable has enough variation, not every country has a significant sample size. Therefore, when analyzing the cancellation rates, we will omit countries with less than 1000 bookings.
```{r}
sort(summary(chotels$country), decreasing = TRUE) [1:13]
ccancelledbycountry <- tapply(chotels$is_canceled, chotels$country, mean, na.rm = TRUE)
sort(ccancelledbycountry, decreasing = TRUE)
```

The countries with the highest number of bookings are PRT, FRA, DEU, GBR, and ESP, and their cancellation rates are about 0.65, 0.20, 0.18, 0.29, and 0.29, respectively.

This variation shows that the cancellation rates are independent from the total number of bookings, so the country of the booking should be considered when calculating the probability of cancellation (if there has been at least 1000 bookings before).
```{r}
data <- data.frame(
  name=c("Portugal (Highest Cancellation)","Germany (Lowest Cancellation)") ,  
  value=c(0.64861111, 0.17619987)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **market_segment**
```{r}
chotels$market_segment <- as.factor(chotels$market_segment)
sort(summary(chotels$market_segment), decreasing = TRUE) 
```

*Variable Validation: though this variable has enough variation, not every market segment has a significant sample size. Therefore, when analyzing the cancellation rates, we will omit market segments with less than 1000 bookings.
```{r}
sort(summary(chotels$market_segment), decreasing = TRUE) [1:5]
ccancelledbymarket <- tapply(chotels$is_canceled, chotels$market_segment, mean, na.rm = TRUE)
sort(ccancelledbymarket, decreasing = TRUE) 
```

As shown in the tables above, the Groups, Offline TA/TO, and Online TA market segments have significantly higher cancellation rates in comparison to other market segments. Given this variation, the hotel should consider the market segment when calculating the probability of cancellation for a booking (if there has been at least 1000 bookings before).
```{r}
data <- data.frame(
  name=c("Groups (Highest Cancellation)","Direct (Lowest Cancellation)") ,  
  value=c(0.6885868, 0.1733136)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```


### **agent**
```{r}
chotels$agent <- as.factor(chotels$agent)
sort(summary(chotels$agent), decreasing = TRUE)
```

*Variable Validation: though this variable has enough variation, not every agent has a significant sample size. Therefore, when analyzing the cancellation rates, we will omit agents with less than 1000 bookings.
```{r}
sort(summary(chotels$agent), decreasing = TRUE) [1:11]
ccancelledbyagent <- tapply(chotels$is_canceled, chotels$agent, mean, na.rm = TRUE)
sort(ccancelledbyagent, decreasing = TRUE)
```

The agents with the highest number of bookings are 9, 1, and 14, and their cancellation rates are about 0.42, 0.73, and 0.18, respectively.

This variation shows that the cancellation rates are independent from the total number of bookings, so the agent of the booking should be considered when calculating the probability of cancellation (if there has been at least 1000 bookings before).
```{r}
data <- data.frame(
  name=c("19 (Highest Cancellation)","28 (Lowest Cancellation)") ,  
  value=c(0.735155514, 0.066026411)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```

### **total_of_special_requests** 
```{r}
chotels$total_of_special_requests <- as.factor(chotels$total_of_special_requests)
sort(summary(chotels$total_of_special_requests), decreasing = TRUE)
```

*Variable Validation: though this variable has enough variation, not every number has a significant sample size. Therefore, when analyzing the cancellation rates, we will omit numbers with less than 1000 bookings.
```{r}
sort(summary(chotels$total_of_special_requests), decreasing = TRUE) [1:4]
ccancelledbyrequest <- tapply(chotels$is_canceled, chotels$total_of_special_requests, mean, na.rm = TRUE)
sort(ccancelledbyrequest, decreasing = TRUE)
```

As shown in the table above, as the number of special requests increase, the cancellation rate decreases significantly, so this should be considered when calculating the probability of cancellation for a booking (if there has been at least 1000 bookings before).
```{r}
data <- data.frame(
  name=c("0 (Highest Cancellation)","5 (Lowest Cancellation)") ,  
  value=c(0.54924203, 0.03846154)
  )
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")

```

## Summary

### Resort Hotel

Below are the variables that have shown statistical significance and correlation to cancellation rates. Thus, these variables should be considered when calculating the probability of cancellation for a booking. With this data, Resort Hotel can develop more accurate predictions of cancellations, which can then be optimized with overbooking.

* lead_time: if under 100 days of lead time, then the probability of cancellation will gradually increase with number of lead time days
* arrival_date_month: highest cancellation rates in August, June, September and lowest cancellation rates in March, November, January 
* stays_in_weekend_nights: if there are under 10 weekend nights booked, the probability of cancellation will gradually increase up to 0.6 at most
* children: consider if children (higher probability of cancellation) or not
* babies: consider if babies or not (higher probability of cancellation)
* country: consider if there has been at least 1000 bookings before
* market_segment: consider if there has been at least 1000 bookings before
* agent: consider if there has been at least 1000 bookings before

### City Hotel

Below are the variables that have shown statistical significance and correlation to cancellation rates. Thus, these variables should be considered when calculating the probability of cancellation for a booking. With this data, City Hotel can develop more accurate predictions of cancellations, which can then be optimized with overbooking.

* lead_time: probability of cancellation will gradually increase with number of lead time days
* stays_in_weekend_nights: if there are under 11 weekend nights booked, the probability of cancellation will gradually increase up to 0.8 at most
* stays_in_week_nights: if there are under 25 week nights booked, the probability of cancellation will gradually increase
* children: consider if children (higher probability of cancellation) or not
* babies: consider if babies or not (higher probability of cancellation)
* country: consider if there has been at least 1000 bookings before
* market_segment: consider if there has been at least 1000 bookings before
* agent: consider if there has been at least 1000 bookings before
* total_of_special_requests: probability of cancellation will gradually decrease with number of special requests but only consider if there has been at least 1000 bookings before
